---
title: "US Wind Turbine"
author: "Alessio Passalacqua"
date: "`r Sys.Date()`"
output: 
  html_document:
    number_sections: false
    toc: true
    toc_float: true
    theme: cosmo

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```



# Motivation

The US Wind turbine database provides onshore & offshore wind turbine locations in the United States, corresponding facility information, and turbine technical specifications.

Database could be download from [here](https://eerscmap.usgs.gov/uswtdb/assets/data/uswtdbCSV.zip)

The authors of database have created a web application called the US wind turbine database viewer available to the [link](https://eerscmap.usgs.gov/uswtdb/viewer/#3/37.25/-96.25).
The viewer lets you visualize inspect and interact with the turbine database through web browser. 


The goal of the analysis is to show the variation of capacity in time and highlight the contribution of each state. 

In first part of the report, the total capacity is showed as a function of time and as a function of states.

In the second part, the evolution of total capacity is showed on map. The final plot represents the total capacity by states in three different years 1998, 2008 and 2018.



# Load library

```{r}
library(readr)       # load data in csv format
library(viridis)     # color palettes
library(dplyr)       # manipulate dataframe
library(sf)          # manipulate shape file
library(ggplot2)     # plot
library(forcats)     # reorder factor levels
library(datasets)    # dataset of US state name
library(leaflet)     # interactive maps
library(DT)          # interactive table
```

# Load data


Load the us wind turbine dataset:
```{r}
usw <- read_csv("uswtdb_v1_1_20180710.csv")
```


# Table

```{r}
usw_dt <- usw %>%
  select(p_year,t_state,t_cap)
datatable(usw_dt, extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
  )
)
```


# Total capacity as function of states and years

Calculate the amount of total capacity added for each year:

```{r}
usw_year <- usw %>%
  filter(t_cap > 0) %>% # remove t_cap negative value (aka missing) 
  filter(p_year >0 ) %>% # remove p_year negative value (aka missing) 
  group_by( p_year) %>%
  summarise(totalcapacity = sum(t_cap)/1000000) #calculate the total capacity in GW
```

Plot total capacity using the `cumsum` function inside ggplot to calculate the cumulative value of total capacity:

```{r}
ggplot(data = usw_year, aes(x = p_year, y = cumsum(totalcapacity))) + geom_line() +
   geom_vline(xintercept = 2006, alpha=0.4,linetype = "dashed" ) +
   annotate("text",x=2003,y=20,label="2006",hjust=0, colour = "gray") +
    labs( x= "", y="Total Capacity [GW]",
       title="US Wind Turbine Total Rated Capacity",
       subtitle = "from 1981 to 2018",
       caption="Source: U.S. Wind Turbine Database")+
    theme_minimal() 
```

The total capacity has a sharp increase after 2006.

Calculate total capacity in each states:


```{r}
usw_states <- usw %>%
  filter(t_cap > 0) %>% # remove t_cap negative value (aka missing) 
  filter(p_year >0 ) %>% # remove p_year negative value (aka missing) 
  group_by( t_state) %>%
  summarise(totalcapacity = sum(t_cap)/1000000) #calculate the total capacity in GW
```

Add state names to `usw_states` dataset by using `state` dataset loaded from `datasets` library:

```{r}
data(state)
states <- cbind(state.name,state.abb)
states <- rbind(states,c("Guam","GU"))
states <- data.frame(rbind(states,c("Puerto Rico","PR")))

# join by abbreviation name column
usw_states <- left_join(usw_states, states,
                          by=c("t_state"="state.abb"))

```

Plot the total capacity as function of states:

```{r,fig.height=8,fig.width=11}
ggplot(usw_states,
       aes(x=totalcapacity, y=fct_reorder(state.name, totalcapacity, desc=TRUE))) +
  geom_segment(  
           aes(x = 0,
           y=fct_reorder(state.name, totalcapacity, desc=TRUE),
           xend = totalcapacity,
           yend = fct_reorder(state.name, totalcapacity, desc=TRUE)),
           color = "gray50") +
           geom_point() +
  labs(x="Total Rated Capacity [GW]", y="", 
       title = "US Wind Turbine",
       subtitle = "2018",
       caption = "Source: U.S. Wind Turbine Database") +
  theme_minimal() +
  # NEW CODE BELOW
  geom_text(aes(label=round(totalcapacity,1)), hjust=-.5) +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_blank(),
        axis.text.x = element_blank())
```

Texas has the higher value of total capacity, about 23 GW.

# Choropleth map of US Wind turbine database

## Load US map shape file

The shape file could be download from the [Census](https://www.census.gov/geo/maps-data/data/tiger-cart-boundary.html) 

Read the shape file of state boundaries:

```{r}
fifty_location <- "static_maps/data/cb_2017_us_state_20m/cb_2017_us_state_20m.shp"
fifty_states <- st_read(fifty_location)
```


## Create dataset for different period

### Wind turbine in 1998

Create a dataframe referning to 1998:
```{r}
usw_tc98 <- usw %>%
  filter(t_cap > 0) %>% # remove t_cap negative (aka missing) 
  filter(p_year >0 & p_year <1999) %>% 
  group_by(t_state) %>%
  summarise(totalcapacity = sum(t_cap)/1000000) #calculate the total capacity in GW
```

Join shape file and usw in 1998:
```{r}
usw_tc98 <- left_join(fifty_states, usw_tc98,
                          by=c("STUSPS"="t_state"))
# Create column year and refered to 1998
usw_tc98$Year <- "1998"
```

### Wind turbine in 2008

Create a dataframe referning to 2008:
```{r}
usw_tc08 <- usw %>%
  filter(t_cap > 0) %>%
  filter(p_year > 0 & p_year < 2009) %>%
  group_by(t_state) %>%
  summarise(totalcapacity = sum(t_cap)/1000000) 
```

Join shape file and usw in 2008:
```{r}
usw_tc08 <- left_join(fifty_states, usw_tc08,
                          by=c("STUSPS"="t_state"))

usw_tc08$Year <- "2008"
```

### Wind turbine in 2018

Create a dataframe referning to 2018:
```{r}
usw_tc18 <- usw %>%
  filter(t_cap > 0) %>%
  filter(p_year > 0) %>%
  group_by(t_state) %>%
  summarise(totalcapacity = sum(t_cap)/1000000)
```

Join shape file and usw in 2018:
```{r}
usw_tc18 <- left_join(fifty_states, usw_tc18,
                          by=c("STUSPS"="t_state"))

usw_tc18$Year <- "2018"
```

## Plot Maps

Aggregate all the dataframes by row:
```{r}
usw_30y <- rbind(usw_tc98,usw_tc08,usw_tc18) 
```

Remove no contiguos states for more compact visualization:
```{r}
usw_30y <- usw_30y %>% 
  filter(NAME!="Hawaii" & NAME!="Alaska" & NAME!="Puerto Rico")
```

Plot maps:

```{r}
ggplot(usw_30y) +
  geom_sf(aes(fill=totalcapacity)) +
  facet_wrap(~Year,ncol=2) +
  labs(title="Wind Turbine Total Rated Capacity of 48 states",
       subtitle = "for R for Journalists class",
       caption="Source: U.S. Wind Turbine Database")+
  scale_fill_viridis(direction=-1, name="GW") +
  scale_color_viridis(direction=-1) +
  theme_void() +
  theme(panel.grid.major = element_line(colour = 'transparent'))
```

# Interactive choropleth map



```{r}
usw_tc18$totalcapacity <- ifelse(is.na(usw_tc18$totalcapacity),0,usw_tc18$totalcapacity) 
pal <- colorNumeric("Oranges", domain=usw_tc18$totalcapacity)
popup_sb <- paste0("Total Capacity: ", as.character(round(usw_tc18$totalcapacity,1)))

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(-98.483330, 38.712046, zoom = 4) %>% 
  addPolygons(data = usw_tc18, 
              fillColor = ~pal(usw_tc18$totalcapacity), 
              fillOpacity = 0.7, 
              weight = 0.2, 
              smoothFactor = 0.2,
              popup = ~popup_sb ) %>%
  addLegend(pal = pal, 
            values = usw_tc18$totalcapacity, 
            position = "bottomright", 
            title = "Total Capacity [GW]")
```


# Connect to Github repository

This lines of code was runned on terminal to connect the local repository to my Github repository

-------------------------

* `git init .` 
* `git add .` 
* `git commit -m "first commit"`
* `git remote add origin https://github.com/alessiopassalacqua/r-journalism.git`
* `git push -u origin master`


-------------------------